### 2018-07-25 pubs/itp-2018/sections/conclusions-and-future-work.tex
(#66) Merging revisions of the paper
* [ paper ] removed many examples and extra content.

* [ paper ] all proofs have been removed.

* Complete first revision to section about Metis

* Review section 4.2

* Review introduction Section 4

* Improve many parts in the text.

Metis section improves in the tptp description.
The example tstp derivation was added with line
numbers.
The explanation of the workflow has changed to
be more explict and also improved for readability.
Remove the numbers in the diagram.
Appendix updated removing the tstp of the small example.

* [ Sec4 ] example of the stri

We presented a proof-reconstruction approach in type theory for the
propositional fragment of the \Metis prover. We provided for each
\Metis inference rule a formal description in type theory following a
syntactical approach. These formalizations are mainly exposed in
Section~\ref{ssec:emulating-inferences}.

We built the \Athena translator tool written in \Haskell
that generates \Agda proof-terms of \Metis derivations.
\Agda files generated by this translator calls theorems from our
\Agda formalizations of the \Metis reasoning~\cite{AgdaProp,AgdaMetis}.

@@

The reconstruction approach in this study was designed to use
only syntactical aspects of the logic.
This decision was at the beginning a drawback
since it demands more detailed proofs, a
description of every transformation or deduction step performed by
the prover, which is rarely included in the output of these programs,
see Section~\ref{ssec:emulating-inferences}.
Nevertheless, we chose that syntactical treatment instead of using
semantics to extend this work towards the support of first-order
logic or other non-classical logics.
For first-order logic, recall satisfiability is undecidable and
its syntactical aspect plays an important role to reconstruct proofs.

This study increased the
trustworthiness of the automatic prover \Metis. Justifying a proof by
a theorem prover has a real significant impact for these automatic
tools.  The reverse engineering task to grasp the prover
reasoning can reveal important issues or bugs in many parts of these
systems (\eg, preprocessing, reasoning, or deduction modules). During
this research, we had the opportunity to contribute to \Metis by
reporting some bugs---see Issues No.~2, No.~4, and commit
\name{8a3f11e} in \Metis' official
repository.\footnote{\url{https://github.com/gilith/metis}.}
Fortunately, all these problems were fixed quickly by Hurd in
\Metis~2.3 (release~20170822).
\end{document}

>>>>

We presented a proof-reconstruction approach in type theory for the
propositional fragment of the \Metis prover. In
Section~\ref{ssec:emulating-inferences}, we provided for each \Metis inference
rule a formal description in type theory following a syntactical approach
type-checked in the proof-assistant \Agda in~\cite{AgdaProp,AgdaMetis}. With
this formalization, we were able to type-check \Agda proof-terms for
\TSTP derivations generated by \Metis. The \Agda proof-terms were generated by
the \Athena translator~\cite{Athena}, a tool written in \Haskell.

@@

Our approach to reconstruct the proofs for \Metis \TSTP derivations only used
syntactical aspects of the logic. We chose that syntactical treatment instead of
using semantics to extend this work towards the support of first-order logic or
other non-classical logics. For instance, let us recall that
for first-order logic, satisfiability is undecidable and its syntactical aspect
plays an important role to reconstruct its proofs.

Lastly, we strongly believe that by justifying proofs by theorem provers, we
help to increase their trustworthiness. In this work, we increased the
trustworthiness for the automatic prover \Metis. The reverse engineering to
grasp the reasoning of a theorem prover can help to reveal issues in these
systems maybe unkwon even for their own developers. During this research, we had
the opportunity to contribute to \Metis by reporting some bugs---see Issues
No.~2, No.~4, and commit \name{8a3f11e} in \Metis' official
repository.\footnote{\url{https://github.com/gilith/metis}.} Fortunately, all
these problems were fixed quickly by Hurd in \Metis~2.3 (release~20170822).
\end{document}

### 2018-07-25 pubs/itp-2018/sections/proof-reconstruction.tex
Add the current output from athena for the article's problem
Nice box color for equations about functions
First attempt to remove complexity measure in sec4
(#66) Merging revisions of the paper
* [ paper ] removed many examples and extra content.

* [ paper ] all proofs have been removed.

* Complete first revision to section about Metis

* Review section 4.2

* Review introduction Section 4

* Improve many parts in the text.

Metis section improves in the tptp description.
The example tstp derivation was added with line
numbers.
The explanation of the workflow has changed to
be more explict a

\begin{myexamplenum}
\label{ex:inference-rule-pattern}
Let \texttt{metisRule} be a \Metis inference rule in the \TSTP derivations.
To reconstruct this \texttt{metisRule}, we define in the type theory
a function preserving the name of the original rule. This function should follow
the following pattern.

\begin{equation*}
  \begin{aligned}
  &\hspace{.495mm}\fun{metisRule} : \Source → \Target → \Prop\\
  &\begin{array}{ll}
  \fun{metisRule}~φ~ψ\ &=
      \begin{cases}
      ψ, &\text{if the conclusion }ψ\text{ can be}\text{ derived by applying certain inference}\\
         &\text{rules to the premise }φ;\\
      φ, &\text{otherwise;}
      \end{cases}
  \end{array}
  \end{aligned}
\end{equation*}

@@

We have finished the formalization of every inference rule in
a \Metis derivation, we are able to justify step-by-step any proof
for a problem in propositional logic.
For instance, we tested successfully the translation by \Athena
jointly with the \Agda formalizations of the rules mentioned above
with more than eighty representative theorems in propositional
logic. An interested reader can test the
problems~\cite{Prieto-Cubides2017} in the \Athena tool
repository~\cite{Athena}.

>>>>

\begin{myexamplenum}
\label{ex:inference-rule-pattern}

Let \texttt{metisRule} be a \Metis inference rule in the \TSTP derivations. To
reconstruct this \texttt{metisRule}, we define in the type theory a function
preserving the name of the original rule. This function should follow the
following pattern.

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation*}
  \begin{aligned}
  &\hspace{.495mm}\fun{metisRule} : \Source → \Target → \Prop\\
  &\begin{array}{ll}
  \fun{metisRule}~φ~ψ\ &=
      \begin{cases}
      ψ, &\text{if the conclusion }ψ\text{ can be}\text{ derived by applying certain inference}\\
         &\text{rules to the premise }φ;\\
      φ, &\text{otherwise.}
      \end{cases}
  \end{array}
  \end{aligned}
\end{empheq}

@@

The above theorem finishes the formalization of the \Metis inference rules. At
this point, we are able to justify step-by-step any proof in propositional logic
generated by \Metis. Actually, we tested successfully the translation by \Athena
jointly with the \Agda formalizations of this section against  more than eighty
representative theorems in propositional logic. The interested reader can
reproduce the testing of all problems from~\cite{Prieto-Cubides2017} with the
sources of the \Athena tool repository~\cite{Athena}.

%% -- TODO : Yo creo que aqui deberia ir el output de Athena...

### 2018-07-25 pubs/itp-2018/sections/proof-reconstruction/strip.tex
(#66) Merging revisions of the paper
* [ paper ] removed many examples and extra content.

* [ paper ] all proofs have been removed.

* Complete first revision to section about Metis

* Review section 4.2

* Review introduction Section 4

* Improve many parts in the text.

Metis section improves in the tptp description.
The example tstp derivation was added with line
numbers.
The explanation of the workflow has changed to
be more explict and also improved for readability.
Remove the numbers in the diagram.
Appendix updated removing the tstp of the small example.

* [ Sec4 ] example of the stri

\begin{notation}
$\Bound$ is a synonym for $\Nat$ type to refer to a natural number that constrains
the number of recursive calls for the function that use it in its type.
\end{notation}

>>>>

\begin{notation}
$\Bound$ is a synonym for $\Nat$ type to refer to a natural number that constrains
the number of recursive calls done by a function. We use this type mostly in
the type of the function.
\end{notation}

### 2018-07-25 pubs/itp-2018/sections/related-work.tex
(#66) Merging revisions of the paper
* [ paper ] removed many examples and extra content.

* [ paper ] all proofs have been removed.

* Complete first revision to section about Metis

* Review section 4.2

* Review introduction Section 4

* Improve many parts in the text.

Metis section improves in the tptp description.
The example tstp derivation was added with line
numbers.
The explanation of the workflow has changed to
be more explict and also improved for readability.
Remove the numbers in the diagram.
Appendix updated removing the tstp of the small example.

* [ Sec4 ] example of the stri

\citeauthor{Kanso2012} in~\cite{Kanso2012,kanso2016light} presents a
proof-reconstruction in \Agda for propositional logic. Its tool support
proof-reconstruction for \name{EProver} and \name{Z3} ATPs following a similar work-flow as we presented in Section~\ref{ssec:workflow}.
Nonetheless, its approach
employs semantics for logic equivalences. We have avoided the use of
propositions meanings towards a future work to support other logics where a
syntactical approach plays an important role~(for an example of such logics, we
refer the reader to \cite{Agudelo-Agudelo2017}).
\citeauthor{foster2011integrating}~\cite{foster2011integrating} describe  the
proof-reconstruction in \Agda for \name{Waldmeister}~\cite{hillenbrand1997}, a
prover for pure equational logic. As far as we know, no other
proof-reconstruction has been carried out neither in \Agda nor with \Metis prover.

Another important proof-assistant in type theory is \name{Coq}~\cite{coqteam}.
We found the \name{SMTCoq}~\cite{armand2011,Ekici2017} tool which provides a
certified checker for proof witness coming from the \SMT solver
\name{veriT}~\cite{bouton2009} and adds a new tactic named verit, that calls
\name{veriT} on any \name{Coq} goal. Also for \name{Coq}, given a fixed but
arbitrary first-order signature, \citeauthor{Bezem2002} in \cite{Bezem2002}
transform a proof produced by the first-order automatic theorem prover
\name{Bliksem}~\cite{deNivelle2003} in a \name{Coq} proof-term.

There are some successful attempts using proof-assistants for classical logic
instead of type theory.
Let us mention some representative of such tools. This description is
mainly based on~\citeauthor{Sicard-Ramirez2016}~\cite{Sicard-Ramirez2016}.

>>>>

\citeauthor{Kanso2012} in~\cite{Kanso2012,kanso2016light} shows a
proof-reconstruction in \Agda using semantics for logic equivalences for
derivations in propositional logic generated by \name{EProver} and \name{Z3}. In
\cite{foster2011integrating}~\citeauthor{foster2011integrating}~describe another
proof-reconstruction in \Agda but for pure equational logic of derivations
generated by \name{Waldmeister}~\cite{hillenbrand1997}. No other
proof-reconstruction has been carried out neither in \Agda nor with \Metis
prover as far as we know.

In \name{Coq}~\cite{coqteam}, other proof-assistant closes to \Agda, we found
the \name{SMTCoq}~\cite{armand2011,Ekici2017} tool which provides a certified
checker for proof witness coming from the \SMT solver
\name{veriT}~\cite{bouton2009} and adds a new tactic named verit, that calls
\name{veriT} on any \name{Coq} goal. \citeauthor{Bezem2002} in~\cite{Bezem2002}
transform a proof produced by the first-order automatic theorem prover
\name{Bliksem}~\cite{deNivelle2003} in a \name{Coq} proof-term given a fixed but
arbitrary first-order signature.

There are other successful attempts to reconstruct proofs from \ATPs but
using proof-assistants for classical logic instead of type theory.
Let us mention some representative tools.
